%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Medium Length Professional CV
% LaTeX Template
% Version 3.0 (December 17, 2022)
%
% This template originates from:
% https://www.LaTeXTemplates.com
%
% Author:
% Vel (vel@latextemplates.com)
%
% Original author:
% Trey Hunner (http://www.treyhunner.com/)
%
% License:
% CC BY-NC-SA 4.0 (https://creativecommons.org/licenses/by-nc-sa/4.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[
	%a4paper, % Uncomment for A4 paper size (default is US letter)
	11pt, % Default font size, can use 10pt, 11pt or 12pt
]{resume} % Use the resume class

\usepackage[T1]{fontenc}
\usepackage{textcomp}
\usepackage{mathspec}
\setmainfont{EB Garamond}
\setmathsfont(Latin){EB Garamond}
\usepackage{hyperref}
\hypersetup{colorlinks=true}

%------------------------------------------------

\name{Jasper Ty} % Your name to appear at the top

% You can use the \address command up to 3 times for 3 different addresses or pieces of contact information
% Any new lines (\\) you use in the \address commands will be converted to symbols, so each address will appear as a single line.

\address{jasperty2@gmail.com \\ \href{https://jasperty.net}{jasperty.net} } % Contact information

%----------------------------------------------------------------------------------------


\begin{document}


\begin{rSection}{Education}
	
    \textbf{Drexel University} 
    \hfill \textit{Philadelphia, USA} \\
    Ph.D in Mathematics
    \hfill 2024-current \\
    Thesis advisor: \href{https://www.math.drexel.edu/~jblasiak/}{Jonah Blasiak}

    \textbf{Drexel University} \hfill \textit{Philadelphia, USA} \\
    B.S. in Computer Engineering
    \hfill 2019-2024

\end{rSection}


\begin{rSection}{Research interests}

    \textbf{Algebraic combinatorics} and representation theory. 

\end{rSection}


\begin{rSection}{Preprints and publications}

    \iffalse
    \textbf{Noncommutative key polynomials for Demazure atom positivity} \\
    Preprint.

    \textbf{Cauchy identities for products under Dyck paths} \\
    Preprint.
    
    \textbf{Flagged LLT polynomials are slide positive} \\
    Preprint.

    \textbf{Raising operator formulas for nonsymmetric plethysm} \\
    Preprint.
\fi
    Still cooking, stay tuned!

\end{rSection}


\begin{rSection}{Selected talks and presentations}

    \section*{Research}

    \textbf{\href{https://darijgrinberg.gitlab.io/cage/abstracts/2025-fall-ty.html}{Cauchy identities for products under Dyck paths}}
    \hfill Dec. 4, 2025 \\
    \href{https://darijgrinberg.gitlab.io/cage/index.html}{CAGE (Combinatorics, Algebra, and Geometry) seminar}

    \textbf{The Lascoux kernel and Demazure atom positivity}
    \hfill Nov. 16, 2025 \\
    BUGCAT (Binghamton University Graduate Combinatorics, Algebra, and Topology) conference

    \textbf{\href{https://www.math.drexel.edu/~tolya/ty\%2010.17.2025.pdf}{The Lascoux kernel and Demazure atom positivity}}
    \hfill Oct. 17, 2025 \\
    \href{https://www.math.drexel.edu/~tolya/analysis.html}{Drexel Analysis seminar}

    \section*{Exposition}

    \textbf{Laplace and $z$-transforms}
    \hfill March. 6, 2026 \\
    Drexel SIAM student chapter seminar

    \textbf{Minesweeper: Logic, Computation, ...Physics?} 
    \hfill Feb. 2, 2026 \\
    Penn Pizza seminar


    \textbf{Integer compositions}
    \hfill Nov. 13, 2025 \\
    Drexel SIAM student chapter seminar

    \textbf{Noncommutative Gröbner bases and Bergman's diamond lemma}
    \hfill Apr. 30, 2025 \\
    Drexel SIAM student chapter seminar

    \textbf{Schubert polynomials} 
    \hfill Jan. 27, 2025 \\
    Drexel SIAM student chapter seminar

    \textbf{Minesweeper: Logic, Computation, ...Physics!?} 
    \hfill Oct. 21, 2024 \\
    Drexel SIAM student chapter seminar


\end{rSection}


\begin{rSection}{Teaching}

    Teaching assistant for \textbf{MATH-200 (Multivariate Calculus)} at Drexel University.
    \hfill Spring 2025-2026

    Teaching assistant for \textbf{MATH-121 (Calculus 1)} at Drexel University.
    \hfill Fall 2025-2026

\end{rSection}


\begin{rSection}{Selected courses}

	\begin{tabular}{@{} >{\bfseries}l @{\hspace{6ex}} l @{}}
        MATH-530 & \href{https://darijgrinberg.gitlab.io/t/25sg/index.html}{Graph Theory} \\
        MATH-531 & \href{https://darijgrinberg.gitlab.io/t/24w/index.html}{Algebraic Combinatorics} \\
		MATH-640 & Functional Analysis \\
		MATH-660 & Lie Groups and Lie Algebras \\
		MATH-670 & Methods of Optimization \\
		MATH-T880 & Positive Polynomials and Sums of Squares \\
		MATH-T880 & Topics in Operator Theory \\
	\end{tabular}
    

\end{rSection}


\begin{rSection}{skills}

	\begin{tabular}{@{} >{\bfseries}l @{\hspace{6ex}} l @{}}
		Software & SageMath, Rust, C, C++, Python, MATLAB, Haskell \\
		Languages & English, French, Waray, Tagalog \\
	\end{tabular}

\end{rSection}



\end{document}

